% !TEX TS-program = pdflatexmk
\documentclass[11pt]{article}
\usepackage[margin=.8in]{geometry}
\usepackage{amsmath,amssymb,amsthm, latexsym, mathrsfs, pdfsync, multicol,
fancybox, fancyhdr,
graphicx, enumerate,
subfig, tikz, pgfplots,array}

%\singlespacing
\def\RR{{\mathbb R}}
\def\NN{{\mathbb N}}
\def\ZZ{{\mathbb Z}}
\def\QQ{{\mathbb Q}}
\def\CC{{\mathbb C}}
\def\bc{\begin{center}}
\def\ec{\end{center}}
\def\be{\begin{enumerate}}
\def\ee{\end{enumerate}}
\def\bi{\begin{itemize}}
\def\ei{\end{itemize}}
\def\t{\times}
\newcommand{\ol}[1]{\overline{#1}}
\newcommand{\oimp}[1]{\overset{#1}{\Longleftrightarrow}}
\newcommand{\bv}[1]{\ensuremath{ \mathbf{\vec{#1}}} }
\renewcommand{\d}{\displaystyle}
\newcommand{\blank}[1]{\rule{#1}{0.75pt}}

\usetikzlibrary{calc}

\lhead{\sc{Math 316 Hist. of Math.}}
\chead{\large \sc Midterm I} 
\rhead{\sc Spring 2023}
\cfoot{}
\pagestyle{fancy}
%
\begin{document}
\thispagestyle{fancy}

\vspace*{-0.3in}

\bc Part I \ec

This part is written without notes or aids of any kind. It is worth 25 points out of 100 total points. \\

Below is a list of eight mathematicians or mathematical documents. Give a detailed description of \textbf{five of the eight}.  A complete description will include approximate dates, location, and mathematical significance. \\

Once you have completed Part I and turn it in, you will be given Part II. You cannot return to Part I once it has been turned in.\\

\be
\item Rhind Papyrus
\vfill
\item Moscow Papyrus
\vfill
\item Plimpton 322
\vfill
\newpage
\item Thales of Miletus
\vfill
\item Pythagoras of Samos
\vfill
\item Euclid
\vfill
\item Theon of Smyrna
\vfill
\item Eudoxus of Cnidos
\vfill
\ee
\newpage 
Sample Answers:

(a) Isaac Newton\\


\noindent (OK answer. 2/5 points)\\
(1600's, England) Founder of Calculus.\\


\noindent (good answer. 4/5 points)\\
(late 1600 - early 1700, England) Along with Gottfried Leibniz, Newton is considered one of two  
mathematicians given credit for developing Calculus. He authored The Principia.\\


\noindent (excellent answer. 5/5 points)\\
(1650-1730, England) Along with Gottfried Leibniz, Newton is considered one of two  
mathematicians given credit for developing Calculus because both realized the tangent line problem and the area under the curve problem were inverse problems. He authored The Principia which formulated his Laws of Motion that with his Calculus provided a simple explanation for planetary motion. \\


\noindent Sentences that are worth 0 points\\

Newton was a famous mathematician who was very influential. His mathematics is still taught today. He authored many manuscripts. He was a professor at Cambridge. He lived during the plague outbreak. He was influenced by Descartes and Galileo.\\


\noindent What I am looking for

\bi
\item Reasonably precise dates
\item Reasonably precise location
\item Two specific, math-related examples of its/their significance.
\ei
\end{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%END
%%%%%%%%%%%%%%%%%%%%%%%%%%


 