% !TEX TS-program = pdflatexmk
\documentclass[12pt]{article}

% Layout.
\usepackage[top=1in, bottom=0.75in, left=1in, right=1in, headheight=1in, headsep=6pt]{geometry}

% Fonts.
\usepackage{mathptmx}
\usepackage[scaled=0.86]{helvet}
\renewcommand{\emph}[1]{\textsf{\textbf{#1}}}

% Misc packages.
\usepackage{amsmath,amssymb,latexsym}
\usepackage{graphicx,hyperref}
\usepackage{array}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{tabularx,colortbl}
\usepackage{enumitem}

\usepackage{fancyhdr}
\pagestyle{fancy} 
\lhead{\large\sf\textbf{MATH 316: History of Math}}
\rhead{\large\sf\textbf{Ticket 17: Sections 6.1-6.2}}

\begin{document}
Jill's Solutions
\begin{enumerate}
\item The term ``the dark ages" references what period in what place?\\
Our text dates this as 400 AD to 1000 AD in Europe. This is sometimes called the Early Middle Ages. The start date is roughly when the Roman Empire fell in Europe. The ending date of 1000 AD is (I believe) about when Western Europe starts to rebound from invasions from the north and east.
\item During this period, what motivated the first systematic attempts at educating people? What was the emphasis of this education?\\
Initially,  the early Christian church required clergy to have some modest education. So monasteries became places to learn to read and write. Later, as Charlemagne consolidated power and built a nation-state, educated people were needed to run governmental operations. The focus was on scripture. Astronomy and mathematics was abandoned in favor of grammar and rhetoric.
\item When does the Prophet Mohammed die and when does the Islamic empire reach northern Spain?\\
632 and 732
 \item Summarize in your own words the paragraph at the bottom of page 272 that continues onto the next page. Do you agree with the last sentence?\\
As the Islamic Empire spread west, Islamic scholars had the opportunity to collect, read, translate, and expand upon manuscripts the Christian Roman Empire had ignored. So the scholarship of Euclid, Archimedes, and Ptolemy came to Europe by way of Arab-speaking scholars, not Latin speaking ones. Preserving what was lost is significant. But it's a little hard to be convinced that Ptolemy's or Euclid's summary and additions to early work was somehow more or better than analogous work of Arab mathematicians such as Al-Khwarizmi...
\item Provide some biographical details about Leonardo of Pisa\\
born 1175 in Pisa Italy. Known as Fibonacci. The famous sequence is named after him because of an example in one of his texts. Educated in North Africa and widely traveled in the Mediterranean as his father was involved in trade. Here he learned about the Hindu-Arabic system and in 1202 published a text called Liber Abaci or Book of Counting.
 \end{enumerate}

\end{document}